Evaluation

### 2018-02-13 evaluation.Rmd
Turn duplicates test back on, typos in evaluation

@@

new_moon_file = portalr::FullPath('PortalData/Rodents/moon_dates.csv', '~')
new_moons = read.csv(new_moon_file)

@@

## How have the models recently?

This graph shows the forecast error of some recent forecasts for the most recent sampling dates. This is shown for all model types in the columns and several common rodent species (and total rodents) in the rows. These forecasts are for the entire site. 
```{r recent_eval, echo=FALSE, message=TRUE, warning=FALSE,, fig.width=15, fig.height=15}
# The length of time to look backward
prior_new_moons = 12

most_recent_newmoon = max(new_moons$newmoonnumber)
earliest_newmoon_to_graph = most_recent_newmoon - prior_new_moons

observation_data = species_abundance %>%
  bind_rows(total_abundance) %>%
  filter(newmoonnumber %in% earliest_newmoon_to_graph:most_recent_newmoon)

#Get all the recent forecasts
forecast_data = compile_forecasts(use_hindcasts = FALSE) %>%
  filter(newmoonnumber %in% earliest_newmoon_to_graph:most_recent_newmoon)

@@

forecast_data = forecast_data %>%
  left_join(forecast_dates_to_keep, by=c('date','initial_newmoon')) %>%
  filter(keep=='yes') %>%
  select(-keep)
##################################

forecast_errors = forecast_data %>%
  left_join(observation_data, by=c('species','level','currency','newmoonnumber')) %>% 
  mutate(rmse = sqrt((estimate - actual)^2))

@@

###################################
forecast_date_labels = forecast_errors %>%
  group_by(date, model, species, censusdate) %>%
  summarize(y_position = mean(rmse)) %>%
  ungroup() %>%
  distinct() %>%
  filter(!is.na(censusdate)) %>%
  group_by(date, model, species) %>%
  filter(as.Date(censusdate) == min(as.Date(censusdate))) %>%
  ungroup() %>%
  distinct() %>%
  mutate(label_text = paste0('Issued: ',date))

####################################
# Setup a nice palette for when prior issue dates become greater than 9
getPalette = colorRampPalette(RColorBrewer::brewer.pal(9, "Set1"))
set.seed(2)
large_color_palette = sample(getPalette(40))
#####################################
ggplot(forecast_errors, aes(x=censusdate, y=rmse, group=as.character(date), color=as.character(date))) +
  geom_point(size=4)+
  geom_line(size=1.5) +
  # label individual issue dates. This is easier to look at if there 
  # gets to be a lot of them. Must install ggrepel and remove the legend if this is turned on.
  #geom_label_repel(data=forecast_date_labels, aes(label=label_text, y=y_position),
  #                 fill='grey90',size=4,nudge_y = 40, force=5,
  #                 segment.alpha = 0.3, segment.size = 0.8) +
  scale_color_brewer(palette='Dark2') +
  #scale_color_manual(values=large_color_palette) +
  facet_grid(species~model) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=0.9, debug = FALSE),
        axis.text = element_text(size=12),
        axis.title = element_text(size=20),
        strip.text = element_text(size=12),
        legend.text = element_text(size=15),
        legend.title = element_text(size=18)) +
  labs(x='Sample Date',y='RMSE', color='Forecast Issue\nDate')

>>>>

models_of_interest = c('AutoArima','Ensemble','ESSS','nbGARCH','pevGARCH')

@@

new_moon_file = portalr::FullPath('PortalData/Rodents/moon_dates.csv', '~')
new_moons = read.csv(new_moon_file)
new_moons$censusdate = as.Date(new_moons$censusdate)

@@

## How have the models done recently?

This graph shows the forecast error of some recent forecasts for the most recent sampling dates. This is shown for all model types in the columns and several common rodent species (and total rodents) in the rows. These forecasts are for the entire site. 
```{r recent_eval, echo=FALSE, message=TRUE, warning=FALSE,, fig.width=15, fig.height=15}
###############################################
#Get all the recent forecasts
forecast_data = compile_forecasts(use_hindcasts = FALSE) %>%
  filter(model %in% models_of_interest)

@@

forecast_data = forecast_data %>%
  left_join(forecast_dates_to_keep, by=c('date','initial_newmoon')) %>%
  filter(keep=='yes') %>%
  select(-keep)

##################################
#Combine forecasts and oservations
forecast_errors = forecast_data %>%
  left_join(observation_data, by=c('species','level','currency','newmoonnumber')) %>% 
  mutate(rmse = sqrt((estimate - actual)^2))

@@

#############################################################
# Displaying the data from many forecasts in the past can create a mess of a plot. see #218.
# So here three forecasts are chosen from 1, 6, and 12 newmoons into the past. 
# Sometimes forecasts data is missing for a particular initial_newmoon, the recursive function here
# attempts to find another one from the same time pluts/minus 2 month period. This ensures 3 forecasts
# are always shown.

# If there is no data for initial_newmoon 500, this will check nearby dates in this order: 499,501,498,402
update_initial_newmoon = function(potential_initial_newmoon, attempt=1){
  if(attempt==6){
    return(NA)
  }
  lead_time_data = forecast_errors %>%
    filter(initial_newmoon == potential_initial_newmoon)

  # Alternate steps of -1, +2, -3, +4
  next_try_step = ifelse(attempt%%2 == 0, attempt, attempt * -1)

  # no data to show? then try one month back
  if(nrow(lead_time_data)==0){
    #print(paste('Try ',attempt, ', Nothing with ',potential_initial_newmoon,' trying ',potential_initial_newmoon+next_try_step))
    return(update_initial_newmoon(potential_initial_newmoon+next_try_step, attempt+1))
  } else {
    return(potential_initial_newmoon)
  }

}
# Adjust this to add more than 3 forecast dates, or change the lead times
lead_times_to_show = c(1,6,12)
most_recent_observed_newmoon = max(new_moons$newmoonnumber)
initial_newmoons_to_keep = most_recent_observed_newmoon - lead_times_to_show

initial_newmoons_to_keep = sapply(initial_newmoons_to_keep, update_initial_newmoon)

forecast_errors = forecast_errors %>%
  filter(initial_newmoon %in% initial_newmoons_to_keep)

#####################################
ggplot(forecast_errors, aes(x=censusdate, y=rmse, group=as.character(date), color=as.character(date))) +
  geom_point(size=4)+
  geom_line(size=1.5) +
  scale_color_brewer(palette='Dark2') +
  scale_x_date(breaks = unique(forecast_errors$censusdate)) +
  facet_grid(species~model, scales = 'free_y') + 
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=0.99, debug = FALSE),
        axis.text = element_text(size=12),
        axis.title = element_text(size=20),
        strip.text = element_text(size=12),
        legend.text = element_text(size=15),
        legend.title = element_text(size=18)) +
  labs(x='Sample Date',y='RMSE', color='Forecast Issue\nDate')
